A large leading tail of atmospheric escape in the HAT-P-67 b system

### 2023-05-25 paper/paper1/ms.tex
Decimate the exospheric grazers subsection

\begin{document}
\shorttitle{Runaway Mass Loss in HAT-P-67b}
\shortauthors{Gully-Santiago et al.}

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/pwinds_data_overlay.png}
    \caption{Simulated 1D Parker wind model of Helium absorption in \hatpb.  The spectrum was generated with a mass loss rate of $10^{13}$ g/s, exosphere gas temperature of 9100 K, and with an XUV luminosity $L_x/L_\mathrm{bol}=10^{-5}$ intermediate between those shown in Figure \ref{fig:XUV}.  The lines have about the same equivalent width, while additional unmodeled velocity dispersion of the gas can explain differences between the line shapes of the model and data.  Overall the 1D models may be too simplistic to represent the inherently 3D structure of the outflow.}
    \label{fig:pwinds}
\end{figure}

@@

We measured a large extent of Helium escape along the arc of the orbit, but we necessarily can place only coarse constraints on the vertical extent---in the direction perpendicular to the orbital path, $\pm z$.  We estimate the detectable vertical extent must be at least a few stellar radii, such that we still could have detected Helium escape in a hypothetical HAT-P-67b-like system even if it were non-transiting, in a ``near miss'' configuration.  We propose a new category of semi-transiting planet, dubbed ``exospheric grazers'', in which the planet does not produce a detectable white-light transit depth, but \emph{does} produce measurable \emph{line}-based exospheric absorption.  This category appears to have been neglected due to the assumption that exospheres were only easily detectable at several planetary radii, and therefore fractions of a stellar radius.  While large tails have been seen previously in Ly$\alpha$, the scarcity of UV resources prohibited searches of this kind.  The discovery of large Helium tails in HAT-P-32b (Zhang et al. 2023) and now HAT-P-67b suggest that the identification of these exospheric grazers may be achievable with current instrumentation and possibly existing archival data from near-IR RV-planet searches.

The configuration would consist of a HAT-P-67b-like planet undergoing profound mass loss, with a jet-like column of material spouting from the poles, lofted there through gas dynamics, or otherwise confined by bow-shock redirection.  A process for identifying strong candidates could select close-in planets or planet candidates discovered from non-transiting techniques, such as RVs or TTVs.  The best candidates would be planets with $M\sin{i}$, orbital periods, and insolations consistent with a hot Saturn mass.  A follow-up program could be devised to acquire in-transit spectroscopy to search for \ion{He}{1} 10830~\AA, or other spectral tracers of mass loss.

A few benefits of this planet category arise.  First, they would expand the sample of planets for which the planetary winds could be measured, by relaxing the narrow geometrical transit probability of $1-10^\circ$ to something closer to $\pm30^\circ$, depending on the mass loss and gas dynamics.  In particular, the smallest stars such as white dwarfs and M-dwarfs could see the search space for exospheres increase by $10-100\times$.  Second, the range of viewing angles would serve as a natural experiment for constraining the vertical extent of outflows in the first place.  The typical bow shock 3D morphology could be deduced from a sample of otherwise-similar planets with a range of orbital inclinations.  Non-detections would therefore be valuable, especially in cases where $\sin{i}$ could be constrained independently.  In practice these additional constraints are rarely possible, so ensemble statistical methods may have to be employed.  Finally, for some promising planet candidates, a distinctive helium signal could validate the planet, assuming the signal could be disambiguated from stellar variability.  That disambiguation may be feasible since the velocity structure and phasing of stellar and planetary Helium signals diverge due to stellar wind confinement and the launching from a predictable planetary rest frame.  The elusive category of young planets could make ideal targets, since their large radii and high intrinsic temperatures and high incident XUV radiation may yield large mass loss rates if they have recently migrated to an interior location through disk migration.  This category is especially valuable if discovered through near-IR RV searches, since abundant multi-epoch spectroscopy may already contain the requisite stellar variability baselines.

>>>>

\begin{document}
\shorttitle{Runaway Mass Loss in the HAT-P-67 system}
\shortauthors{Gully-Santiago et al.}

@@

\begin{figure}
    \includegraphics[width=\linewidth]{figures/pwinds_data_overlay.png}
    \caption{Simulated 1D Parker wind model of Helium absorption in \hatpb.  The spectrum was generated with a mass loss rate of $10^{13}$ g/s, exosphere gas temperature of 9100 K, and with an XUV luminosity $L_x/L_\mathrm{bol}=10^{-5}$ intermediate between those shown in Figure \ref{fig:XUV}.  The \texttt{p-winds} model yields the \ion{He}{1} 10830 triplet lines (colored dashed lines), with a cumulative feature in black possessing about the same equivalent width as the data, shown as the in-transit mean.  Additional unmodeled velocity dispersion of the gas can explain differences between the line shapes of the model and data.  Overall the 1D models may be too simplistic to represent the inherently 3D structure of the outflow.}
    \label{fig:pwinds}
\end{figure}

@@

We measured a large extent of Helium escape along the arc of the orbit, but we necessarily can place only coarse constraints on the vertical extent---in the direction perpendicular to the orbital path, $\pm z$.  We estimate the detectable vertical extent must be at least a few stellar radii, such that we still could have detected Helium escape in a hypothetical HAT-P-67b-like system even if it were non-transiting, in a ``near miss'' configuration.  We propose a new category of semi-transiting planet, dubbed ``exospheric grazers'', in which the planet does not produce a detectable white-light transit depth, but \emph{does} produce measurable \emph{line}-based exospheric absorption.  This category appears to have been neglected due to the assumption that exospheres were only easily detectable at several planetary radii, and therefore fractions of a stellar radius.  While large tails have been seen previously in Ly$\alpha$, the scarcity of UV resources prohibited searches of this kind.  The discovery of large Helium tails in HAT-P-32b (Zhang et al. 2023) and now HAT-P-67b suggest that the identification of these exospheric grazers may be achievable with current instrumentation and possibly existing archival data from near-IR RV-planet searches.  Non-transiting, strongly irradiated planets with well-constrained orbits may make ideal targets for detection of this phenomenon.
